---
title: "Homework 8"
author: "Amber Clark - SDS 315 UT Austin"
output:
  html_document:
      toc: true
      toc_float: true
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

------------------------------------------------------------------------


[Github Link](https://github.com/ambertclark/SDS-315.git) 

```{r message=FALSE, echo=FALSE, warning=FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)
```

# **1. Question One**
```{r message=FALSE, echo=FALSE, warning=FALSE}
creatinine <- read.csv("creatinine.csv")
# linear model for creatinine data
model <- lm(creatclear ~ age, data = creatinine)

# A) predict creatinine clearance for 55 year old
pred_55 <- predict(model, newdata = data.frame(age = 55))

# B) How does creatinine rate change with age?
coef_age <- coef(model)["age"]

# C) Whose creatinine rate is healthier for their age?

pred_40 <- predict(model, newdata = data.frame(age = 40))
pred_60 <- predict(model, newdata = data.frame(age = 60))


```

A). The creatinine clearance rate that we should expect for a 55-year-old is 113.72 mL/minute. I determined this by inputting the age 55 into the linear regression equation: 

creatclear = 147.81 - 0.6198 x age.

B). The coefficient for the linear regression model is -0.619159 meaning that as age increases, the creatinine clearance rate is expected to decrease by -0.62 mL/min on average. I determined this by interpreting the coefficient of the linear regression model which is -0.619159.

C). The residual for a 40-year-old with a clearance of 135 is approximately +11.98 and the residual for a 60-year-old with a clearance of 112 is approximately +1.38, meaning that the 40-year old is healthier for their age because their creatinine clearance is about 11.98 mL/min above expected, while the 60-year-old is only about 1.38 mL/min above the expected. I determined this by finding the residuals (actual - expected) and subtracting the given actual creatinine clearance (135 and 112) by the creatinine clearance predicted by plugging in the ages in the model (123.02 and 110.62).

For 40-year-old
Expected = 147.81 - 0.6198 x 40 = 123.02
Residual = 135 - 123.02 = +11.98

For 60-year-old
Expected = 147.81 - 0.6198 x 60 = 110.62
Residual = 112 - 110.62 = +1.38


# **2. Question Two**
```{r message=FALSE, echo=FALSE}

covid <- read.csv("covid.csv")

# fit linear model to 'deaths' and 'days_since_first_death'
italy <- covid %>% filter(country == "Italy", deaths > 0)
spain <- covid %>% filter(country == "Spain", deaths > 0)

# 1). estimate for italy
model_italy <- lm(log(deaths) ~ days_since_first_death, data = italy)


# finding growth rate and doubling time for italy

growth_italy <- coef(model_italy)[2]

doubling_italy <- log(2) / growth_italy


# bootstrapping growth rate and doubling time for italy

boot_italy <- do(1000) * {
  sample <- resample(italy)
  model <- lm(log(deaths) ~ days_since_first_death, data = sample)
  growth <- coef(model)[2]
  tibble(
    growth = growth,
    doubling = log(2) / growth
    
  )
}

# 95% CI for growth rate and doubling time for italy
CI_italy_growth <- quantile(boot_italy$growth, c(0.025, 0.975))
CI_italy_doubling <- quantile(boot_italy$doubling, c(0.025, 0.975))

# 2). estimate for spain
model_spain <- lm(log(deaths) ~ days_since_first_death, data = spain)

# finding growth rate and doubling time for spain

growth_spain <- coef(model_spain)[2]
doubling_spain <- log(2) / growth_spain

# bootstrapping growth rate and doubling time for spain

boot_spain <- do(1000) * {
  sample <- resample(spain)
  model <- lm(log(deaths) ~ days_since_first_death, data = sample)
  growth <- coef(model)[2]
  tibble(
    growth = growth,
    doubling = log(2) / growth
  )
}

# 95% CI for growth rate and doubling time for spain

CI_spain_growth <- quantile(boot_spain$growth, c(0.025, 0.975))
CI_spain_doubling <- quantile(boot_spain$doubling, c(0.025, 0.975))

# line graph

ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1.2) +
  labs(
    title = "Daily COVID Deaths",
    x = "Days Since First Death",
    y = "Daily Reported Deaths"
  ) + theme_minimal() + scale_color_manual(values = c("Italy" = "pink", "Spain" = "skyblue"))
```

1). An estimated growth rate and doubling time with 95% bootstrapped confidence intervals for Italy is 0.183 (0.158, 0.209) and 3.8 (3.3, 4.4) respectively.

2). An estimated growth rate and doubling time with 95% bootstrapped confidence intervals for Spain is 0.276 (0.234, 0.317) and 2.5 (2.2, 3.0) respectively.

3). The line graph above shows the reported daily deaths over time.

# **3. Question Three**

```{r message=FALSE, echo=FALSE}
milk <- read.csv("milk.csv")

# removing rows with 0 to avoid log(0)
milk <- milk %>% filter(sales > 0)

# create variables changed to log
milk <- milk %>%
  mutate(log_price = log(price),
         log_sales = log(sales))

# log-log linear model
model <- lm(log_sales ~ log_price, data = milk)

# estimate for elasticity
elasticity <- coef(model)[2]

# bootstrapping elasticity
boot_elasticity <- do(1000) * {
  sample <- resample(milk)
  mod <- lm(log_sales ~ log_price, data = sample)
  tibble(elasticity = coef(mod)[2])
}

# 95% CI for elasticity
ci <- quantile(boot_elasticity$elasticity, c(0.025, 0.975))
```

1). I estimated the price elasticity to be -1.619 with a 95% confidence interval (-1.772, -1.466). I  estimated the price elasticity of demand by fitting a log-log linear model and using 1000 resampled bootstraps of the data:

log(Q) = log(K) + βlog(P)

